\RequirePackage[l2tabu, orthodox]{nag}
\documentclass[fontsize=11pt]{scrartcl}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{bm}
\usepackage[a4paper]{geometry}
\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{siunitx}
\usepackage{xcolor}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex,bookmarks,colorlinks]{hyperref}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{newtxtext}
\usepackage[varg]{newtxmath}
\usepackage{booktabs}
\usepackage{datetime}
\usepackage{setspace}
\usepackage{enumitem}
\usepackage{cleveref}
\usepackage[numbers]{natbib}
\bibliographystyle{abbrvnat}
\setlist{itemsep=-3pt}
\title{Nondimentionalization}
\date{\shortdate\today \, \ampmtime }
\author{seto}
\begin{document}
\maketitle

\subsection*{Force balance equation}

The following force balance equation is considered here;
\begin{equation}
 \bm{F}_{\mathrm{H}} +  \bm{F}_{\mathrm{R}}  = 0.
\end{equation}
%
The hydrodynamic interaction is
assumed to be linear form of velocities.
%
\begin{equation}
 \bm{F}_{\mathrm{H}} = - \bm{R}\cdot\bm{U}.\label{110733_3Jun15}
\end{equation}

Once we set unit of force,
force balance equations can be nondimensionalized.
%
Then, we also need to select
relation between force unit and velocity unit.
%
Since our force balance equations include
Stokes  hydrodynamic interactions \eqref{110733_3Jun15},
the simplest way is use of Stokes drag formula
for any force units:
\begin{equation}
 F^{\ast} = 6 \pi \eta_{0} a U^{\ast}.\label{220656_3Jun15}
\end{equation}
%
Thanks to this,
the hydrodynamic interaction is nondimensionalized as follows,
\begin{equation}
 \bm{F}_{\mathrm{H}}/F^{\ast}
  = -(\bm{R}/6\pi\eta_0 a) \cdot (\bm{U}/U^{\ast})
     = -\bm{R}' \cdot (\bm{U}/U^{\ast}),
\end{equation}
%
and the resistance matrix in our codes
is the identity matrix
in the dilute limit: $\bm{R}'(\phi\to 0) = \bm{I}$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%




\subsection*{Representative quantities}

Particle radius is commonly considered
as the representative length in
our particle simulations: $L^{\ast} = a$.




\subsubsection*{From shear rate}

One way to give the representative particle velocity
in a sheared suspension is
\begin{equation}
U_{\mathrm{H}}^{\ast} = L^{\ast} \dot{\gamma}.
\end{equation}
%
%
The typical hydrodynamic force is given
by the typical velocity $U_{\mathrm{H}}^{\ast} $
and Stokes-drag form;
\begin{equation}
  F_{\mathrm{H}}^{\ast} = 6 \pi \eta_0 a U_{\mathrm{H}}^{\ast}.
\end{equation}
Here, we indicate nondimensionalized variables
in the hydrodynamic unit by using `tilde',
i.e.,
$\tilde{\bm{F}}  = \bm{F}/ F_{\mathrm{H}}^{\ast} $,
etc.
%

[note]
Since $F_{\mathrm{H}}^{\ast}$ is a function of shear rate $\dot{\gamma}$,
it is not constant in stress controlled simulations
and zero if system is not sheared.


\subsubsection*{From repulsive force}

The repulsive force $\bm{F}_{\mathrm{R}}$ 
is a function of particle-to-particle distance.
%
We take the amplitude $F_{\mathrm{R}}^{\ast}$
to represent its typical strength:
%
\begin{equation}
 \bm{F}_{\mathrm{R}} =  F_{\mathrm{R}}^{\ast}
\hat{\bm{F}}_{\mathrm{R}},
\end{equation}
where `hat' indicates nondimensionalized variables
based on this repulsive force.
%
The corresponding velocity can also be
given via the Stokes-drag form;
\begin{equation}
 U^{\ast}_{\mathrm{R}}
  \equiv \frac{F_{\mathrm{R}}^{\ast}}{6\pi\eta_0 a}.
\end{equation}
%

\subsection*{Hydrodynamic (or shear-rate) unit}

In the hydrodynamic unit, the nondimensionalized
force balance equation is
\begin{equation}
 \frac{\bm{F}_{\mathrm{H}}}{F^{\ast}_{\mathrm{H}}}
  +
  \frac{\bm{F}_{\mathrm{R}}}{F^{\ast}_{\mathrm{H}}}  = 0
  \quad
  \Longrightarrow 
    \quad
 - \bm{R}' \cdot \tilde{\bm{U}} + \tilde{\bm{F}}_{\mathrm{R}} = 0,
\end{equation}
where
$\bm{R}' \equiv \bm{R} / 6 \pi \eta_0 a $.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The origin of shear-rate dependence
becomes clear if the repulsive force
is expressed with the typical value $F_{\mathrm{R}}^{\ast}$;
%
\begin{equation}
 - \tilde{\bm{R}} \cdot \tilde{\bm{U}} +
\frac{1}{\alpha}  \hat{\bm{F}}_{\mathrm{R}} = 0.
\end{equation}
%
This $\alpha$ indicates the competition between two representative forces:
\begin{equation}
 \alpha \equiv
\frac{\bm{F}_{\mathrm{H}}^{\ast}}{\bm{F}_{\mathrm{R}}^{\ast}}
=
  \frac{\dot{\gamma}}{F_{\mathrm{R}}^{\ast}/6\pi\eta_0 a^2}\label{102734_31May15}.
\end{equation}


The unit of time is
\begin{equation}
 t_{\mathrm{H}}^{\ast} \equiv \frac{L^{\ast}}{U_{\mathrm{H}}^{\ast}}
  = \frac{1}{\dot{\gamma}}.
\end{equation}
Therefore, the nondenominational time $\tilde{t}$
is shear strain (when shear rate is constant):
\begin{equation} 
 \tilde{t} \equiv \frac{t}{t_{\mathrm{H}}^{\ast}} = t \dot{\gamma}
   = \gamma.
\end{equation}
%
The nondimensionalized shear rate is
\begin{equation}
 \tilde{\dot{\gamma}}
  =
  \frac{ \dot{\gamma}}{1/t_{\mathrm{H}}^{\ast}}
   = 1.
\end{equation}


[note]
We call this unit system ``hydrodynamic''.
But, the essential origin for this unit system
is the way to give typical velocity,
i.e., $U^{\ast} = a \dot{\gamma}$.
%
``Shear-rate'' unit might be proper name.


\subsection*{Repulsive force unit}

In the repulsive unit,
the nondimensionalized
force balance equation is
\begin{equation}
 \frac{\bm{F}_{\mathrm{H}}}{F^{\ast}_{\mathrm{R}}}
  +
  \frac{\bm{F}_{\mathrm{R}}}{F^{\ast}_{\mathrm{R}}}  = 0
      \quad
  \Longrightarrow
    \quad
    - \bm{R}' \cdot \hat{\bm{U}} +
    \hat{\bm{F}}_{\mathrm{R}} = 0.
\end{equation}
The time unit is
\begin{equation}
 t_{\mathrm{R}}^{\ast}
  \equiv \frac{L^{\ast}}{U_{\mathrm{R}}^{\ast}}
  = \frac{6 \pi \eta_0 a^2}{F_{\mathrm{R}}^{\ast}}.
\end{equation}
%
The nondimensionalized shear rate is
\begin{equation}
 \hat{\dot{\gamma}} =
  \dot{\gamma} t_{\mathrm{R}}^{\ast}
  = \alpha.
\end{equation}



In rate-controlled simulations,
the relation between
the nondimensionalized time
$\hat{t} \equiv t /t_{\mathrm{R}}^{\ast}$
and shear strain $\gamma$ is
\begin{equation}
 \gamma
  \equiv \dot{\gamma} t 
  = \dot{\gamma} t_{\mathrm{R}}^{\ast} \hat{t} 
   = 
\frac{F_{\mathrm{H}}^{\ast}}{F_{\mathrm{R}}^{\ast}}
\hat{t} = \alpha \hat{t},
\end{equation}
where $\alpha$ is defined in \eqref{102734_31May15}.

In stress controlled simulations,
the strain is simply given by
\begin{equation}
 \gamma
 = \int^{t} \dot{\gamma} dt
 = \int^{\hat{t}} \alpha d\hat{t}.
\end{equation}

\subsection*{Simulation data}

In both ways of nondimensionalization,
we can unify the output data with ones nondimensionalized by repulsive force.

\begin{center}
  \begin{tabular}{lccccc}
    &   {\bf Output} & {\bf Output in H.U.} &  {\bf Calc in H.U.}
    & {\bf Output in R.U.} &  {\bf Calc in R.U.}
    \\
    {\bf strain} & $\gamma $ & $ \tilde{t}$
    & & $\gamma $
    \\
    {\bf relative viscosity}   &  $ \eta_{\mathrm{r}}$ &
    $6 \pi \tilde{\sigma}$ &  &
    $6 \pi \hat{\sigma}/\alpha$  & \\
    {\bf time}  & $\hat{t}$ & $ \tilde{t}/\alpha$
    &  $\tilde{t} $ & $\hat{t}$ & $\hat{t}$\\
    {\bf shear rate}& $\hat{\dot{\gamma}}$ & $\alpha$ & $\tilde{\dot{\gamma}} = 1$
    & $\hat{\dot{\gamma}}$ & $\hat{\dot{\gamma}}$  \\
    {\bf shear stress} & $\hat{\sigma}$
       & $\alpha\tilde{\sigma}$  & $\tilde{\sigma} $  &
      $\hat{\sigma}$ &  $\hat{\sigma}$ 
  \end{tabular} 
\end{center}


\subsection*{Note}

This choice of relating
velocity scale and force scale \eqref{220656_3Jun15}
is specific.
%
This is why
we see numerical factor $6\pi$ at several places.

This may generate a little confusion
on nondimensionalization of stresses.
%
In viscous suspensions,
the representative value of stress
can be given by the solvent viscosity $\eta_0$
and shear rate.
\begin{equation}
 \sigma^{\ast} = \eta_0 \dot{\gamma}^{\ast}.
\end{equation}
%
We need to give a typical shear rate in the system.
%
If a repulsive force plays
the critical role on the shear rate dependence,
we can set the corresponding shear rate $\dot{\gamma}^{\ast}_{\mathrm{R}}$:
\begin{equation}
 F_{\mathrm{R}}^{\ast} =
  6 \pi \eta_0 a^2 \dot{\gamma}^{\ast}_{\mathrm{R}}
\end{equation}
So, the unit of stress can be given as follows,
\begin{align}
 \sigma^{\ast}_{\mathrm{R}}
 &= \eta_0 \dot{\gamma}^{\ast}_{\mathrm{R}} \notag \\
  &=  F_{\mathrm{R}}^{\ast}/6\pi a^2
\end{align}
Thus, the numerical factor $6\pi$
comes from the way to relate force and velocity.





\end{document}